---
title: "20200222_EQ-Vas Bsln to 6mo"
author: "Emma Lee"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(qwraps2)
library(janitor)
library(ggplot2)
```

# EQ VAS #

#### Raw data is sourced from the following locations: 
####     1. Bsln EQ VAS from APICS-01, OSEAR Only, APICS-C --> APICS-01 Redcap database 
####     2. 6mo EQ Vas from APICS-01, OSEAR Only --> APICS-01 Access database
####     3. 6mo EQ Vas from APICS-C --> APICS-C Access database 


## Read in the Baseline data
```{r read_bsln, include=FALSE}

bsln_data <- read_csv(here::here("Data", "20210311_OSEAR APICS01 Bsln.csv"))

# Pull EQ VAS column
    vars <- c("id", "osear_1")
    eqvas_bsln <- bsln_data[vars]
    
    eqvas_bsln <- eqvas_bsln %>% 
      filter(is.na(osear_1) == FALSE)

# Rename 
    colnames(eqvas_bsln)[2] <- "eqvas_dc"

    
# Clean up
rm(vars)
rm(bsln_data)
```

## Read in the 6mo follow up data 
```{r read_6mo, include=FALSE}

f_up_data <- read_csv(here::here("Data", "20210311_APICS01 OSEAR 6mo Data.csv"))

# Pull EQ VAS column 
    vars2 <- c("p1_subid", "p6_eqol_health")
    eqvas_6mo <- f_up_data[vars2] 

# Rename 
    colnames(eqvas_6mo)[1] <- "id"
    colnames(eqvas_6mo)[2] <- "eqvas_6mo"
    
    
# Clean up
rm(vars2)
rm(f_up_data)
```

## Check for duplicates and merge datasets 
```{r merge, include=FALSE}

# BSLN Missings
    pt_list <- read_csv(here::here("Data", "20210204_OSEAR Enrolled ID List.csv")) %>% 
      mutate(id = participant_id) %>%
      select(id)

    bsln_list <- as.data.frame(unique(eqvas_bsln$id))
    colnames(bsln_list) <- "id"

    setdiff(pt_list, bsln_list)                                 # Missing those 3 IMC pts 2011, 2014, 2015
    
# BSLN Duplicated
    eqvas_bsln <- distinct(eqvas_bsln)
    
    bsln_dup <- eqvas_bsln %>% 
      filter(duplicated(id) == TRUE)                            # Yasss no duplicates


# 6MO Missings 
    
    fup_list <- as.data.frame(unique(eqvas_6mo$id))
    colnames(fup_list) <- "id"
    
    setdiff(pt_list, fup_list)
    
# 6MO Duplicates
    
    eqvas_6mo <- distinct(eqvas_6mo)                             # Yasss no duplicates again


eqvas_data <- eqvas_bsln %>% 
  left_join(eqvas_6mo)


rm(pt_list)
rm(bsln_list)
rm(bsln_dup)
rm(fup_list)

```

## Summary of data
```{r summary}

# Counts 
    sum(!is.na(eqvas_data$eqvas_dc))        # 177 baseline 
    sum(!is.na(eqvas_data$eqvas_6mo))       # 125 complete @ 6mo
    sum(is.na(eqvas_data$eqvas_6mo))        # 52 missing @ 6mo

# Baseline summary stats 

    summary(eqvas_data$eqvas_dc)

    hist(eqvas_data$eqvas_dc, 
         main = "Expected Health 6mo from Discharge", 
         xlab = "% of Full Health", 
         xlim = c(0, 100), 
         ylim = c(0, 60), 
         breaks = 7)


# 6mo summary stats 
    
    summary(eqvas_data$eqvas_6mo)
    
    hist(eqvas_data$eqvas_6mo, 
         main = "Perceived Health at 6mo from Discharge", 
         xlab = "% of Full Health", 
         xlim = c(0, 100), 
         ylim = c(0, 60), 
         breaks = 10)

    
# Boxplot comparison 

    boxplot(eqvas_data[,-1], 
            horizontal = TRUE, 
            main = "Expected vs. Perceived Health at 6mo Post-Discharge", 
            xlab = "% of Full Health")
    
    
```
## Baseline to 6mo changes into DF
```{r change, echo=FALSE}

# Calculate differences + label up/down/same 

    change_data <- eqvas_data %>% 
      mutate(diff = (eqvas_data$eqvas_6mo - eqvas_data$eqvas_dc)) 
    
    change_data <- change_data %>% 
      mutate(change_label = ifelse(diff > 0, "Perceived > Expected", NA)) %>% 
      mutate(change_label = ifelse(diff < 0, "Perceived < Expected", change_label)) %>% 
      mutate(change_label = ifelse(diff == 0, "Perceived = Expected", change_label))
    
    change_data <- change_data %>% 
      select(id, diff, change_label) %>% 
      filter(!is.na(diff))


```


## Summary of changes
```{r change_summary}

# Counts of changes
    
    sum(!is.na(change_data$diff))                # 125 complete @ 6mo
    nrow(change_data[change_data$diff > 0,])     # 27 perceived > expected 
    nrow(change_data[change_data$diff == 0,])    # 25 perceived = expected
    nrow(change_data[change_data$diff < 0,])     # 75 perceived < expected
    
    nrow(change_data[change_data$diff > 8, ])    # 16 perceived > expected by more than 8 points 
    nrow(change_data[change_data$diff < -8,])    # 64 perceived < expected by more than 8 points 
    
    nrow(change_data[change_data$diff > 10, ])   # 11 perceived > expected by more than 10 points 
    nrow(change_data[change_data$diff < -10,])   # 51 perceived < expected by more than 10 points
    
    nrow(change_data[change_data$diff > 20, ])   #  2 perceived > expected by more than 20 points 
    nrow(change_data[change_data$diff < -20,])   # 24 perceived < expected by more than 20 points

    
    hist(change_data$diff, 
         main = "Distribution of changes from Expected to Perceived Health", 
         xlab = "% Perceived Health - % Expected Health", 
         xlim = c(-100, 100), 
         ylim = c(0, 50), 
         breaks = 12)
    

```


## Spaghetti plot
```{r spaghetti, echo=FALSE}

# Prepare data with a nice massage 

    # Add a timepoint column for bsln data, labelled "value"
    eqvas_bsln <- eqvas_bsln %>% 
      mutate(timepoint = "Expected at D/C") 
    colnames(eqvas_bsln)[2] <- "value"
    
    # Add a timepoint column for 6mo data, labelled "value"
    eqvas_6mo <- eqvas_6mo %>% 
      mutate(timepoint = "Perceived at 6mo")
    colnames(eqvas_6mo)[2] <- "value"
    
    # Rbind bsln and 6mo data 
    data <- rbind(eqvas_bsln, eqvas_6mo)
    data <- arrange(data, id)
    
    # Merge change 
    data <- data %>% 
      left_join(change_data)
    
    # Select complete cases 
    data <- data[complete.cases(data), ]
    

# Base R version of the spaghetti plot 
#    interaction.plot(x.factor = data$timepoint, 
#                     trace.factor = data$id, 
#                     response = data$value, 
#                     ylab = "% of Full Health",
#                     xlab = "Timepoint",
#                     lty = 1, 
#                     lwd = 1, 
#                     legend = FALSE
#                     ) 
    # meh

    
# Trying out ggplot2 version 
    
    #data2 <- filter(data, !is.na(data$eqvas_6mo))

    noodles <- data %>%
      ggplot(aes(x = timepoint, y = value, color = change_label)) + 
      geom_line(aes(group = id), 
                position = position_jitter(w=0.02, h=0)) + 
      labs(title = "Expected vs. Perceived Health 6mo Post-Discharge", 
           x = "", 
           y = "% of Best Possible Health", 
           color = "Legend") + 
      scale_color_manual(values=c("black", "green", "red")) + 
      theme_bw()

    print(noodles)
    # better


```








