---
title: "ADL IADL WHOQOL"
author: "Emma Lee"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(janitor)
```

# BSLN to 6mo ADL/IADL + WHOQOL-BREF Scores # 
#### Quick data pull for grant work 
####    Baseline data sourced from cleaned output from 20210223_Bsln Cleaning.Rmd 
####    6mo Data sourced from Access database 

# Read in raw data 
```{r read-in, include=FALSE}

# Baseline Data 
    bsln_data <- read_csv(here::here("Data", "20210305_OSEAR Baseline Data.csv"))
    bsln_codebook <- read_csv(here::here("Data", "20210306_OSEAR Baseline Codebook.csv"))
    colnames(bsln_data) <- colnames(bsln_codebook)

# 6 month data 
    fup_data <- read_csv(here::here("Data", "20210311_APICS01 OSEAR 6mo Data.csv"))
    fup_codebook <- read_csv(here::here("Data", "20210311_APICS01 OSEAR 6mo Codebook.csv"))
    colnames(fup_data) <- colnames(fup_codebook)


```

# Pick out baseline expectations for ADLs/IADLs
```{r bsln_dls, include=FALSE}

vars1 <- c("id", str_subset(colnames(bsln_data), "expect_", negate=FALSE))

expectations <- bsln_data %>% 
  select(vars1) %>% 
  select(-expect_groom)


rm(vars1)
rm(bsln_data, bsln_codebook)
```

# Pick out the 6mo ADLs and IADLs 
```{r 6mo_dls, include=FALSE}

vars2 <- c("id", str_subset(colnames(fup_data), "adl6_", negate=FALSE))

outcome <- fup_data %>% 
  select(vars2) %>% 
  select(-adl6_done, -adl6_admin, -adl6_contin, -iadl6_done, -iadl6_admin, -iadl6_laundry)


rm(vars2)

```

# Determine whether expectations were met or met/exceeded, count # independent ADL/IADLs
```{r expectations_met, include=FALSE}

data <- left_join(expectations, outcome)

pt_list <- unique(data[, 1])

## NOTE ON PAIRING EXPECTATION VS. OUTCOME VARIABLES: 
##    Baseline includes "Grooming - such as brushing teeth or combing your hair" which is not asked in 6mo ADL/IADLs
##    6mo includes "Continence" (ADL) and "Laundry" (IADL) which are not asked about at Baseline 
## In total there are 12 Expectation-Outcome variable pairings 


# Were the expectations for ADL/IADL at 6mo met exactly? 
#     Expectation and outcome must be exact matches 
#     I.e. Expected: No (0),  Outcome: Yes (1) = Expectation NOT met 

met <- data %>% 
  mutate(expmet_feed = ifelse((expect_feed == adl6_feed), 1, 0)) %>% 
  mutate(expmet_bath = ifelse((expect_bath == adl6_bath), 1, 0)) %>% 
  mutate(expmet_dress = ifelse((expect_dress == adl6_dress), 1, 0)) %>% 
  mutate(expmet_toil = ifelse((expect_toil == adl6_toil), 1, 0)) %>% 
  mutate(expmet_transf = ifelse((expect_transf == adl6_transf), 1, 0))

met <- met %>% 
  mutate(expmet_housekeep = ifelse((expect_housekeep == 1 & iadl6_housekeep == 1) | (expect_housekeep == 0 & iadl6_housekeep > 1), 1, 0)) %>% 
  mutate(expmet_fin = ifelse((expect_fin == 1 & iadl6_fin == 1) | (expect_fin == 0 & iadl6_fin > 1), 1, 0)) %>% 
  mutate(expmet_transport = ifelse((expect_transport == 1 & iadl6_transport == 1) | (expect_transport == 0 & iadl6_transport > 1), 1, 0)) %>% 
  mutate(expmet_foodprep = ifelse((expect_foodprep == 1 & iadl6_foodprep == 1) | (expect_foodprep == 0 & iadl6_foodprep > 1), 1, 0)) %>% 
  mutate(expmet_shop = ifelse((expect_shop == 1 & iadl6_shop == 1) | (expect_shop == 0 & iadl6_shop > 1), 1, 0)) %>% 
  mutate(expmet_meds = ifelse((expect_meds == 1 & iadl6_meds == 1) | (expect_meds == 0 & iadl6_meds > 1), 1, 0)) %>% 
  mutate(expmet_phone = ifelse((expect_phone == 1 & iadl6_phone == 1) | (expect_phone == 0 & iadl6_phone > 1), 1, 0)) 
  
met <- met %>% 
  mutate(expmet_score = rowSums(met[, (str_detect(colnames(met), "expmet_") == TRUE)])) %>% 
  mutate(allmet = ifelse(expmet_score == 12, "Yes", "No")) 
  

# Were the expectations for ADL/IADL at 6mo met OR exceeded? 
#     Outcome can match or exceed expectation 
#     I.e. Expected: No (0), Outcome: Yes (1) = Expectation met/exceeded 

met_exceed <- data %>% 
  mutate(exceed_feed = ifelse((expect_feed == adl6_feed) | (adl6_feed == 1), 1, 0)) %>% 
  mutate(exceed_bath = ifelse((expect_bath == adl6_bath) | (adl6_bath == 1), 1, 0)) %>% 
  mutate(exceed_dress = ifelse((expect_dress == adl6_dress) | (adl6_dress == 1), 1, 0)) %>%
  mutate(exceed_toil = ifelse((expect_toil == adl6_toil) | (adl6_toil == 1), 1, 0)) %>%
  mutate(exceed_transf = ifelse((expect_transf == adl6_transf) | (adl6_transf == 1), 1, 0)) 
  
met_exceed <- met_exceed %>% 
  mutate(exceed_housekeep = ifelse((expect_housekeep == 0) | (iadl6_housekeep == 1), 1, 0)) %>%
  mutate(exceed_fin = ifelse((expect_fin == 0) | (iadl6_fin == 1), 1, 0)) %>%
  mutate(exceed_transport = ifelse((expect_transport == 0) | (iadl6_transport == 1), 1, 0)) %>%
  mutate(exceed_foodprep = ifelse((expect_foodprep == 0) | (iadl6_foodprep == 1), 1, 0)) %>%
  mutate(exceed_shop = ifelse((expect_shop == 0) | (iadl6_shop == 1), 1, 0)) %>%
  mutate(exceed_meds = ifelse((expect_meds == 0) | (iadl6_meds == 1), 1, 0)) %>%
  mutate(exceed_phone = ifelse((expect_phone == 0) | (iadl6_phone == 1), 1, 0)) 

met_exceed <- met_exceed %>% 
  mutate(exceed_score = rowSums(met_exceed[, (str_detect(colnames(met_exceed), "exceed_") == TRUE)])) %>% 
  mutate(all_exceed = ifelse(exceed_score == 12, "Yes", "No"))


# How many ADLs/IADLs can pt do independently at 6mo? Anything other than highest level of independence counts as dependent 

indep <- outcome %>% 
  mutate(iadl6_phone = ifelse(iadl6_phone == 1, 1, 0)) %>% 
  mutate(iadl6_shop = ifelse(iadl6_shop == 1, 1, 0)) %>% 
  mutate(iadl6_foodprep = ifelse(iadl6_foodprep == 1, 1, 0)) %>% 
  mutate(iadl6_housekeep = ifelse(iadl6_housekeep == 1, 1, 0)) %>% 
  mutate(iadl6_transport = ifelse(iadl6_transport == 1, 1, 0)) %>% 
  mutate(iadl6_meds = ifelse(iadl6_meds == 1, 1, 0)) %>% 
  mutate(iadl6_fin = ifelse(iadl6_fin == 1, 1, 0)) 
  
indep <- indep %>% 
  mutate(num_indep_total = rowSums(indep[, str_detect(colnames(indep), "adl6_") == TRUE])) %>%
  mutate(num_indep_iadl = rowSums(indep[, str_detect(colnames(indep), "iadl6_") == TRUE])) %>% 
  mutate(num_indep_adl = (num_indep_total - num_indep_iadl)) 




# Join all them things together 
expectations_met <- list(pt_list, indep, met, met_exceed) %>% reduce(left_join, by = "id") %>% 
  select(id, num_indep_adl, num_indep_iadl, num_indep_total, expmet_score, allmet, exceed_score, all_exceed) 


rm(data, expectations, outcome, met, indep, met_exceed)
```

# WHOQOL-BREF subset 
```{r whoq_subset, include=FALSE}

vars3 <- c("id", str_subset(colnames(fup_data), "OSEAR_6mo1_", negate=FALSE))

whoq_data <- fup_data %>% 
  select(vars3) %>% 
  select(-OSEAR_6mo1_whoqol_done)

colnames(whoq_data) <- c("id", "qol", "satisf", "pain", "med", "enjoy", "mean", "concent", "safe", "phys", "energ", "appear", "mon", "avail", "leis", "around", "sleep", "adl", "work", "self", "relat", "sex", "frenz", "livplace", "health", "transport", "blues")

rm(vars3)

```

# WHOQOL-BREF Scoring
```{r whoq_score, include=FALSE}

# Verify appropriate ranges for each answer 
    range_who <- whoq_data %>% 
      select(-id)

    stopifnot(range(range_who, na.rm = TRUE) == c(1, 5))
    
    rm(range_who)

# Flip reverse-coded questions 
    for(var in c('pain', 'med', 'blues')) {
      whoq_data <- whoq_data %>% 
       dplyr::mutate(!!var := case_when(get(var) == 1 ~ 5, 
                                     get(var) == 2 ~ 4, 
                                     get(var) == 3 ~ 3, 
                                     get(var) == 4 ~ 2, 
                                     get(var) == 5 ~ 1))
    }
   
     
# Domain raw scores: D1 physical health, D2 psychological health, D3 social relationships, D4 environment
#     D1 (6 Qs) = 3 pain, 4 med, 10 energ, 15 around, 16 sleep, 17 adl, 18 work
#     D2 (5 Qs) = 5 enjoy, 6 mean, 7 concent, 11 appear, 19 self, 26 blues
#     D3 (2 Qs) = 20 relat, 21 sex , 22 frenz
#     D4 (6 Qs) = 8 safe, 9 phys, 12 mon, 13 avail, 14 leis, 23 livplace, 24 health, 25 transport

whoq_scores <- whoq_data %>% 
  dplyr::mutate(d1_score = 4*rowMeans(whoq_data[, c('pain', 'med', 'energ', 'around', 'sleep', 'adl', 'work')], na.rm = TRUE)) %>% 
  dplyr::mutate(d2_score = 4*rowMeans(whoq_data[, c('enjoy', 'mean', 'concent', 'appear', 'self', 'blues')], na.rm = TRUE)) %>% 
  dplyr::mutate(d3_score = 4*rowMeans(whoq_data[, c('relat', 'sex', 'frenz')], na.rm = TRUE)) %>% 
  dplyr::mutate(d4_score = 4*rowMeans(whoq_data[, c('safe', 'phys', 'mon', 'avail', 'leis', 'livplace', 'health', 'transport')], na.rm = TRUE))

# Round off those decimals 
whoq_scores <- whoq_scores %>% 
  round(digits=1) %>% 
  select(id, d1_score, d2_score, d3_score, d4_score)

    
rm(var, fup_data, fup_codebook)
```


# Quick export for Alison's grant analysis 
```{r adl_iadl_whoq, include=FALSE}

export <- left_join(expectations_met, whoq_scores)

write_csv(export, "20210315_ADL IADL WHOQ Prelim Export.csv")


```


# Look at WHOQOL data
```{r whoq_summaries, echo = FALSE}

## D1: Physical Health 
summary(whoq_data$d1_score)
hist(whoq_data$d1_score, 
     main = "WHOQOL-BREF D1: Physical Health", 
     xlab = "Score (4-20)", 
     xlim = c(4, 20), 
     ylim = c(0, 50), 
     breaks = 8)

## D2: Psychological Health 
summary(whoq_data$d2_score)
hist(whoq_data$d2_score, 
     main = "WHOQOL-BREF D2: Psychological Health", 
     xlab = "Score (4-20)", 
     xlim = c(4, 20), 
     ylim = c(0, 50), 
     breaks = 5)

## D3: Social Relationships 
summary(whoq_data$d3_score)
hist(whoq_data$d3_score, 
     main = "WHOQOL-BREF D3: Social Relationships", 
     xlab = "Score (4-20)", 
     xlim = c(4, 20), 
     ylim = c(0, 50), 
     breaks = 8)

## D4: Environment
summary(whoq_data$d4_score)
hist(whoq_data$d4_score, 
     main = "WHOQOL-BREF D4: Environment", 
     xlab = "Score (4-20)", 
     xlim = c(4, 20), 
     ylim = c(0, 50), 
     breaks = 8)


```











