---
title: "3_OSEAR_Baseline_Exposures"
author: "Alison E. Turnbull"
date: "05/26/2022"
output: html_document
---
## This script assumes that all code in the files 1_OSEAR_Baseline_6mo_merge.Rmd and 2_OSEAR_Baseline_demographics.Rmd have already been run.
## The codebook for the baseline demographic data cleaned in this script is a file in this project called: "20210306_OSEAR Baseline Codebook.csv"

## This is also where you need to be really careful to have the most utd version of the dplyr package and check that re-coding of factors in the 3rd chunk is
## working properly before you change the "dat" dataframe. 

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(dplyr)
library(janitor)
```


```{r expected health, include=FALSE, echo=FALSE}

#table(is.na(dat$osear_1))                
#dat[which(is.na(dat$osear_1)), "id"]    # Missing for 3: 2011, 2014, 2015  

## Spoke with OACIS LTO team on 3/16/2022.  They sent the following file with missing data on expected health and expected I/ADLs.
temp <- read_csv(here::here("data", "20220316_missing_IMC_exposures.csv"))

## So, for all 3 participants, 2011, 2014, 2015, the VAS is 80, and they expect to do all ADLS, and IADLS indepdently

dat[which(is.na(dat$osear_1)), "osear_1"]<-80

dat<-dplyr::rename(dat, exp_VAS=osear_1)
# quantile(dat$exp_VAS, na.rm = TRUE)
# hist(dat$exp_VAS, na.rm=TRUE)
```

```{r expected I/ADLs, include=FALSE, echo=FALSE}
# dat[which(is.na(dat$osear_2_a)), "id"]    

# I am going to read a list of 13 activities.  For each activity, tell me if you expect that you'll be able to perform it *without help*
# 6 months from now. For each activity say: "yes", "no", or "unsure."
# Note: Yes = 1, No = 0, and Unsure = "2"

exp_ADLs<-dat %>%
            select(starts_with("osear_2_"))  %>%
            rename_with(~str_replace(., "osear_2", "expect"), everything()) %>%
                  mutate(across(everything(), ~factor(.x))) %>%
                          mutate(across(everything(), ~fct_recode(.x, Yes = "1", 
                                                                      No  = "0", 
                                                                      Unsure="2"))) ## This converts all the expected I/ADLs to factors

## Again, look at "temp" from the previous chunk. 
## This variable was missing for 2011, 2014, and 2015.  It was then retrieved on 3/16/2022.  All 3 participants expected to do everything independently.
exp_ADLs<-exp_ADLs %>%
              replace(is.na(.), "Yes")

# colSums(is.na(exp_ADLs))  Yup, no more missings


## Re-naming the 13 variables that capture what people EXPECTED to be able to do in 6 months.
exp_ADLs<-exp_ADLs %>%
  dplyr::rename(exp_feed=expect_a) %>%   #Eating food (someone else can prepare the food)
  dplyr::rename(exp_bath=expect_b) %>%   #Showering or taking a bath
  dplyr::rename(exp_groom=expect_c) %>%  #Grooming - such as brushing teeth or combing your hair
  dplyr::rename(exp_dress=expect_d) %>%  #Getting dressed (putting on clothes)
  dplyr::rename(exp_toil=expect_e) %>%   #Using a toilet
  dplyr::rename(exp_bed=expect_f) %>%    #Moving from a bed to a chair on the other side of the room
  dplyr::rename(exp_clean=expect_g) %>%  #Cleaning your home or living space
  dplyr::rename(exp_money=expect_h) %>%  #Managing money (even if you can't pay all bills)
  dplyr::rename(exp_mobile=expect_i) %>% #Driving a car or taking public transportation 
  dplyr::rename(exp_cook=expect_j) %>%   #Cooking or preparing meals
  dplyr::rename(exp_shop=expect_k) %>%   #Shopping for items you need
  dplyr::rename(exp_meds=expect_l) %>%   #Taking medications in the correct dosage at the correct time
  dplyr::rename(exp_phone=expect_m)      #Using a telephone

#Now we run into the issue of the ADLs and IADLs I asked about at discharge and at follow-up not matching. 
# The 12 that were asked about during BOTH interviews are as follows.
# ADLS:  bathing, dressing, toilet, transfer, feeding
# IADLS: Telephone, shopping, food prep , housekeeping, transport, medications, money


# Calculating how many of the I/ADLs most people expected to be able to do in 6 months.
 n_iadls<-dat %>%
            select(starts_with("osear_2_"))  %>%
              replace(is.na(.), 1) %>%
                select(-osear_2_c) %>%                                            # This drops the grooming question which is not part of the final 12
                  dplyr::mutate(across(everything(), ~ifelse(.x ==1, 1, 0))) %>%
                    rowwise() %>%
                      dplyr::mutate(n_expects=sum(c_across(cols=everything()))) %>%
                          select(n_expects)
 
dat<-bind_cols(dat, exp_ADLs, n_iadls) 
rm(exp_ADLs, n_iadls) 
```

```{r important I/ADLs, include=FALSE, echo=FALSE}
# Now I'm going to read the same list of activities, but this time I want to know how important performing each activity without help is
# for your personal satisfaction with your life in 6 months.  Remember, the question is: 
# "How important is it to me that I can do this activity without help 6 months from now?"


# dat[which(is.na(dat$osear_3_a)), "id"] --> Again, 2011, 2014, and 2015 - the same 3 from Utah are missing.            

imp_ADLs<-dat %>%
            select(starts_with("osear_3_"))  %>%
            rename_with(~str_replace(., "osear_3", "important"), everything()) %>%
                  mutate(across(everything(), ~factor(.x))) %>%
                          mutate(across(everything(), ~fct_recode(.x, Not = "0", 
                                                                      Slightly = "1", 
                                                                      Moderately="2", 
                                                                      Important ="3", 
                                                                      Very     = "4"))) ## Converts importance of I/ADLs to factors.

# Re-naming the 13 variables that capture what people felt was IMPORTANT to be able to do in 6 months.
imp_ADLs<-imp_ADLs %>%
  dplyr::rename(imp_feed=important_a) %>%   #Eating food (someone else can prepare the food)
  dplyr::rename(imp_bath=important_b) %>%   #Showering or taking a bath
  dplyr::rename(imp_groom=important_c) %>%  #Grooming - such as brushing teeth or combing your hair
  dplyr::rename(imp_dress=important_d) %>%  #Getting dressed (putting on clothes)
  dplyr::rename(imp_toil=important_e) %>%   #Using a toilet
  dplyr::rename(imp_bed=important_f) %>%    #Moving from a bed to a chair on the other side of the room
  dplyr::rename(imp_clean=important_g) %>%  #Cleaning your home or living space
  dplyr::rename(imp_money=important_h) %>%  #Managing money (even if you can't pay all bills)
  dplyr::rename(imp_mobile=important_i) %>% #Driving a car or taking public transportation 
  dplyr::rename(imp_cook=important_j) %>%   #Cooking or preparing meals
  dplyr::rename(imp_shop=important_k) %>%   #Shopping for items you need
  dplyr::rename(imp_meds=important_l) %>%   #Taking medications in the correct dosage at the correct time
  dplyr::rename(imp_phone=important_m)      #Using a telephone

# Calculating how many of the I/ADLs most people expected to be able to do in 6 months.
 n_imp<-dat %>%
            select(starts_with("osear_3_"))  %>%
              dplyr::mutate(across(everything(), ~ifelse(.x >2, 1, 0))) %>%
                  rowwise() %>%
                      dplyr::mutate(n_important=sum(c_across(cols=everything()))) %>%
                          select(n_important)

dat<-bind_cols(dat, imp_ADLs, n_imp) 
rm(imp_ADLs, n_imp, temp)  
```

