---
title: "5_Finalizing dataset for analysis"
author: "Alison E. Turnbull"
date: "5/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script assumes that all code in the files:
##            1_OSEAR_Baseline_6mo_merge.Rmd  
##            2_OSEAR_Baseline_demographics.Rmd 
##            3_OSEAR_Baseline_baseline_exposures.Rmd and
##            4_OSEAR_Outcomes.Rmd have already been run.
## The codebook for the baseline demographic data cleaned in this script is a file in this project called: "20210306_OSEAR Baseline Codebook.csv"

## Be really careful to have the most utd version of the dplyr package and check that re-coding of factors is working properly before you change the "dat" dataframe. 

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(janitor)
```


```{r demographics, include=FALSE, echo=FALSE}
demographics<-dat %>%
                select(id, status_6mo, age, sex, race, hispanic, edu, 
                       working, sickleave, retired, disability, unemployed, baseline_employment.unknown,
                          income, adi_dec, adi_perc, 
                            insur_private, insur_Mcaid, insur_Mcare, insur_None,
                            adl_feed, adl_bath, adl_dress, adl_toil, adl_trans, adl_housekeep, adl_money, adl_transport, adl_cook, adl_shop, adl_med, adl_phone, 
                              adm_dx, apache, sx, covid, ards, los, depress, anxiety, cfs,
                                mspss_total, mspss_mean, mspss_so, mspss_fam, mspss_fr,
                                  cd_score,
                                    dc_adl_bath, dc_adl_dress, dc_adl_toil, dc_adl_bed, dc_adl_cont, dc_adl_feed)

demographics<-demographics %>%
                clean_names()

# colSums(is.na(demographics))

```

```{r exposures, include=FALSE, echo=FALSE}
exposures<-dat %>%
    select(id, status_6mo, exp_VAS, 
                exp_feed, exp_bath, exp_dress, exp_toil, exp_bed, exp_clean, exp_money, exp_mobile, exp_cook, exp_shop, exp_meds, exp_phone, n_expects,
                  n_important, imp_feed, imp_bath, imp_dress, imp_toil, imp_bed, imp_clean, imp_money, imp_mobile, imp_cook, imp_shop, imp_meds, imp_phone)

exposures <- exposures %>%
  clean_names()

# colSums(is.na(exposures))
```


```{r outcomes, include=FALSE, echo=FALSE}
outcomes<-dat %>%
      select(id, status_6mo,working_6mo, disability_6mo, employment_6mo.unknown, retired_6mo, sickleave_6mo, unemployed_6mo,
                eqol_5d_code, eqol_6mo_source, eqol_VAS, eq_index, 
                  eqol_activities, eqol_anxiety, eqol_mobility, eqol_pain, eqol_selfcare,
                    eqol_activities_pxy, eqol_anxiety_pxy, eqol_mobility_pxy, eqol_pain_pxy, eqol_selfcare_pxy,
                          whoqol_d1_100, whoqol_d2_100, whoqol_d3_100, whoqol_d4_100, whoqol_d1, whoqol_d2, whoqol_d3, whoqol_d4,
                            whoqol_appearance, whoqol_around, whoqol_available, whoqol_capacity, whoqol_concentrate, whoqol_daily, whoqol_energy, 
                            whoqol_enjoy, whoqol_friends, whoqol_helath, whoqol_leisure, whoqol_meaningful, whoqol_med, whoqol_money, whoqol_negative,
                            whoqol_pain, whoqol_physical, whoqol_place, whoqol_relationships, whoqol_safe, whoqol_sex, whoqol_sleep, 
                            whoqol_transport, whoqol_yourself)

outcomes<-bind_cols(outcomes, iadls_6mo, mspss_6mo)

outcomes<-outcomes %>%
    clean_names()

outcomes<-dplyr::rename(outcomes, whoqol_health=whoqol_helath)

df<-demographics %>%
            left_join(exposures, by=c("id", "status_6mo")) %>%
              left_join(outcomes, by=c("id", "status_6mo"))

#colSums(is.na(outcomes[which(outcomes$status_6mo=="alive"), ]))
 
```

```{r exploring missingness, include=FALSE, echo=FALSE}
df_alive<- df %>%
        filter(status_6mo=="alive")

n_missing<-colSums(is.na(df_alive))
missings=as_tibble(as.list(n_missing))
rm(n_missing)

########################### I/ADL missingness: 
df_alive %>%
  select(starts_with("mo6_")) %>%
      dplyr::summarise(across(everything(), ~sum(is.na(.x))))

df_alive[which(is.na(df_alive$mo6_bathing)), "id"]


######################### EQOL missingness - #3208
# df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_activities)), "id"] 
# df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_mobility)), "id"] 
# df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_anxiety)), "id"]
# df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_pain)), "id"]
# df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_selfcare)), "id"]
# 
# df_alive[which(df_alive$eqol_6mo_source=="Proxy" & is.na(df_alive$eqol_activities_pxy)), "id"] 
# df_alive[which(df_alive$eqol_6mo_source=="Proxy" & is.na(df_alive$eqol_mobility_pxy)), "id"] 
# df_alive[which(df_alive$eqol_6mo_source=="Proxy" & is.na(df_alive$eqol_anxiety_pxy)), "id"]
# df_alive[which(df_alive$eqol_6mo_source=="Proxy" & is.na(df_alive$eqol_pain_pxy)), "id"]
# df_alive[which(df_alive$eqol_6mo_source=="Proxy" & is.na(df_alive$eqol_selfcare_pxy)), "id"]   #3131 is strange, but is NOT missing any data

table(is.na(df_alive$eqol_vas), df_alive$eqol_6mo_source)  # Missing for 2 pts who completed EQ-5D
df_alive[which(df_alive$eqol_6mo_source=="Patient" & is.na(df_alive$eqol_vas)), c("id", "eqol_6mo_source", "eqol_vas")] #3028 and 2020

######################## MSPSS missingness
df_alive %>%
  select(starts_with("mspss_6mo")) %>%
      dplyr::summarise(across(everything(), ~sum(is.na(.x))))

df_alive[which(is.na(df_alive$mspss_6mo_total)), "id"]

###################### WHOQOL_BREF missingness
df_alive %>%
  select(starts_with("whoqol_")) %>%
      dplyr::summarise(across(everything(), ~sum(is.na(.x))))

# df_alive[which(is.na(df_alive$whoqol_appearance)), "id"]
# df_alive[which(is.na(df_alive$whoqol_capacity)), "id"]
# df_alive[which(is.na(df_alive$whoqol_daily)), "id"]
# df_alive[which(is.na(df_alive$whoqol_money)), "id"]
# df_alive[which(is.na(df_alive$whoqol_pain)), "id"]
# df_alive[which(is.na(df_alive$whoqol_yourself)), "id"]

```

```{r clean up, echo=FALSE}
coenrolled<-coenrolled$id
osear_only<-osear_only$id

rm(dat, df_alive, iadls_6mo, missings, mspss_6mo, temp, baseline_IDs, demographics, baselines, exposures, outcomes, pre_covid, coenrolled)

```

