---
title: "5_Finalizing dataset for analysis"
author: "Alison E. Turnbull"
date: "3/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script assumes that all code in the files:
##            1_OSEAR_Baseline_6mo_merge.Rmd  
##            2_OSEAR_Baseline_demographics.Rmd 
##            3_OSEAR_Baseline_baseline_exposures.Rmd and
##            4_OSEAR_Outcomes.Rmd have already been run.
## The codebook for the baseline demographic data cleaned in this script is a file in this project called: "20210306_OSEAR Baseline Codebook.csv"

## Be really careful to have the most utd version of the dplyr package and check that re-coding of factors is working properly before you change the "dat" dataframe. 

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(janitor)
```


```{r demographics, include=FALSE, echo=FALSE}

```

```{r exposures, include=FALSE, echo=FALSE}

```

```{r outcomes, include=FALSE, echo=FALSE}

```
