---
title: "5_Finalizing dataset for analysis"
author: "Alison E. Turnbull"
date: "3/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script assumes that all code in the files:
##            1_OSEAR_Baseline_6mo_merge.Rmd  
##            2_OSEAR_Baseline_demographics.Rmd 
##            3_OSEAR_Baseline_baseline_exposures.Rmd and
##            4_OSEAR_Outcomes.Rmd have already been run.
## The codebook for the baseline demographic data cleaned in this script is a file in this project called: "20210306_OSEAR Baseline Codebook.csv"

## Be really careful to have the most utd version of the dplyr package and check that re-coding of factors is working properly before you change the "dat" dataframe. 

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(janitor)
```


```{r demographics, include=FALSE, echo=FALSE}
demographics<-dat %>%
                select(id, status_6mo, age, sex, race, hispanic, edu, 
                       working, sickleave, retired, disability, unemployed, baseline_employment.unknown,
                          income, adi_dec, adi_perc, 
                            insur_private, insur_Mcaid, insur_Mcare, insur_None,
                              adm_dx, sx, covid, ards, depress, anxiety, cfs,
                                mspss_total, mspss_mean, mspss_so, mspss_fam, mspss_fr,
                                  cd_score,
                                    dc_adl_bath, dc_adl_dress, dc_adl_toil, dc_adl_bed, dc_adl_cont, dc_adl_feed)

demographics<-demographics %>%
                clean_names()

colSums(is.na(demographics))

```

```{r exposures, include=FALSE, echo=FALSE}

```

```{r outcomes, include=FALSE, echo=FALSE}

```
