---
title: "Demographics"
author: "Emma Lee"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(plyr)
library(dplyr)
library(qwraps2)
library(janitor)
options(qwraps2_markup = "markdown")
```

# Demographics #

#### Raw data is sourced from 2 locations 
####     a) Screening data from Redcap "APICS-Screening" 
####         - Age
####         - Gender
####         - Race
####         - Ethnicity
####     b) Baseline data from Redcap "APICS-Data Collection Forms" 
####         - All of the above 
####         - Prior HUS 
####         - Type of Insurance 
####         - Admission date / OSH status 
####         - Admission Dx


## APICS-Screening demographic characteristics table 
```{r screening, include=FALSE }

## This code was used periodically to update the enrollment counts using the bare bones demographics information collected at participant screening ## 

# Read in demographics data from APICS screening Redcap export (age, gender, race, ethnicity)
dem_data_scr <- read_csv(here::here("Data", "20210204_OSEAR Demographics.csv"))

# Split participants along axes of gender and ethnicity 
      #Female & Not Hispanic or Latino 
      dem_f_nhl <- dem_data_scr %>% 
        filter(gender == 2, ethn == 0)
      #Female & Hispanic/Latino 
      dem_f_hl <- dem_data_scr %>% 
        filter(gender == 2, ethn == 1)
      #Female & unknown ethnicity 
      dem_f_unk <- dem_data_scr %>% 
        filter(gender == 2, ethn == 2)
      
      #Now the males 
      dem_m_nhl <- dem_data_scr %>% 
        filter(gender == 1, ethn == 0)
      dem_m_hl <- dem_data_scr %>% 
        filter(gender == 1, ethn == 1)
      dem_m_unk <- dem_data_scr %>% 
        filter(gender == 1, ethn == 2)

# Build a basic Table 1
summary1<- 
      list("Age" = 
             list("median (IQR)" = ~qwraps2::median_iqr(.data$age, digits = 0)), 
           "Gender n (%)" = 
             list("Female"= ~qwraps2::n_perc0(.data$gender == 2), 
                  "Male" = ~qwraps2::n_perc0(.data$gender == 1)), 
           "Race n (%)" = 
             list("White"= ~qwraps2::n_perc0(.data$race == 4),
                  "Black" = ~qwraps2::n_perc0(.data$race == 2), 
                  "Multiracial" = ~qwraps2::n_perc0(.data$race == 6),
                  "Asian" = ~qwraps2::n_perc0(.data$race == 1), 
                  "Unknown" = ~qwraps2::n_perc0(.data$race == 7)),
           "Ethnicity n (%)" =
             list("Hispanic or Latino" = ~qwraps2::n_perc0(.data$ethn == 1), 
                  "Not Hispanic or Latino" = ~qwraps2::n_perc0(.data$ethn == 0), 
                  "Unknown" = ~qwraps2::n_perc0(.data$ethn == 2)))

# Assembling the columns of Table 1 
  tab1<-summary_table(dem_data_scr, summary1) 
  
  tab1_f_nhl<-summary_table(dem_f_nhl, summary1)
  tab1_f_hl<-summary_table(dem_f_hl, summary1)
  tab1_f_unk<-summary_table(dem_f_unk, summary1)
  
  tab1_m_nhl<-summary_table(dem_m_nhl, summary1)
  tab1_m_hl<-summary_table(dem_m_hl, summary1)
  tab1_m_unk<-summary_table(dem_m_unk, summary1)

# Print Table  
  print(tab1,                                   # Modify tab1_* name to view counts for subgroups in table form
      rtitle = "Study Participant Characteristics")


```


## APICS-Data Collection Forms demographic characteristics table
```{r baseline, include=FALSE }

## This code is to be used to generate Table 1 for participants co-enrolled in APICS-01 ##

# Read in entire Baseline csv 
bsln_data_a01 <- 




# Filter for demographic info only 
dem_data_bsln <- data %>% 
  select(id, )
  
# Read in OSEAR enrollment ID list 
osear_ids <- read_csv(here::here("Data", "20210204_OSEAR Enrolled ID List.csv"))

  
summary1<- 
      list("Age" = 
             list("median (IQR)" = ~qwraps2::median_iqr(.data$age, digits = 0)), 
           "Gender n (%)" = 
             list("Female"= ~qwraps2::n_perc0(.data$gender == "Female"), 
                  "Male" = ~qwraps2::n_perc0(.data$gender == "Male")), 
           "Race n (%)" = 
             list("White"= ~qwraps2::n_perc0(.data$race == "White"),
                  "Black" = ~qwraps2::n_perc0(.data$race == "Black"), 
                  "Multiracial" = ~qwraps2::n_perc0(.data$race == "Multiracial"),
                  "Asian" = ~qwraps2::n_perc0(.data$race == "Asian"), 
                  "Unknown" = ~qwraps2::n_perc0(.data$race == "Unknown")),
           "Ethnicity n (%)" =
             list("Hispanic or Latino" = ~qwraps2::n_perc0(.data$ethn == "Yes"), 
                  "Not Hispanic or Latino" = ~qwraps2::n_perc0(.data$ethn == "No"), 
                  "Unknown" = ~qwraps2::n_perc0(.data$ethn == "Unknown")),
            "Education n (%)" = 
             list("8th grade or less" = ~qwraps2::n_perc0(.data$ed == "8th grade or less"), 
                  "Some high school, but did not graduate" = ~qwraps2::n_perc0(.data$ed == "Some high school")
                  "High school graduate or GED" = ~qwraps2::n_perc0(.data$ed == "High school graduate or GED"),
                  "Some college or 2-year degree" = ~qwraps2::n_perc0(.data$ed == "Some college or 2-year degree"), 
                  "4-year college graduate" = ~qwraps2::n_perc0(.data$ed == "4-year degree"), 
                  "More than 4-year college degree" = ~qwraps2::n_perc0(.data$ed == "More than 4-year degree"), 
                  "Not available" = ~qwraps2::n_perc0(.data$ed == "Not available / Unknown")), 
           "ADI Decile" = 
             list("median (IQR)" = ~qwraps::median_iqr(.data$adi_dec, digits = 0)), 
           "ADI Percentile" = 
             list("median (IQR)" = ~qwraps::median_iqr(.data$adi_perc, digits = 0)))
  

```




